---
title: "Class09"
author: "Tim"
format: pdf
---


```{r}
candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```


Q1. How many different candy types are in this dataset?
#85

```{r}
nrow(candy)
```

Q2. How many fruity candy types are in the dataset?
#38



```{r}
library(d)
```



```{r}
table(candy[,2])
table(candy$fruity)
```


Q3. What is your favorite candy in the dataset and what is it’s winpercent value?
#73.09956

```{r}
candy["Milky Way", ]$winpercent
```

Q4. What is the winpercent value for “Kit Kat”?
#76.7686

```{r}
candy["Kit Kat",]$winpercent
```


Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?
#49.6535

```{r}
candy["Tootsie Roll Snack",]$winpercent
```

```{r}
library(dplyr)
candy |>
  filter(rownames(candy) %in% c("Kit Kat", "Milky Way")) |>
  select(winpercent)
```

the %in% operator used to check the intervention of two vectors. 


```{r}
candy |>
  filter(winpercent > 75) |>
  filter(pricepercent < 0.5)
```



```{r}
library("skimr")
skim(candy)
```

Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
#the winpercent data is so high since the data was not scaled 


Q7. What do you think a zero and one represent for the candy$chocolate column?
# they are False = 0 and True = 1, which shows if the individual candy has these items in it or not



Q8. Plot a histogram of winpercent values

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent) +
  geom_histogram(bins=10) +
  theme_bw()
```

```{r}
summary(candy$winpercent)
```


Q9. Is the distribution of winpercent values symmetrical?
#NO

Q10. Is the center of the distribution above or below 50%?
#little above 50

Q11. On average is chocolate candy higher or lower ranked than fruit candy?


```{r}
inds <- as.logical(candy$chocolate)
candy[inds,]$winpercent
```

```{r}
candy |>
  filter(chocolate == 1) |>
  select(winpercent)
```

```{r}
inds <- candy$fruity ==1
fruit.win <- candy[inds,]$winpercent

inds <- as.logical(candy$chocolate)
choc.win <- candy[inds,]$winpercent
```


```{r}
summary(choc.win)
summary(fruit.win)
```

Q12. Is this difference statistically significant?
#yes becasue the P-value is below 0.5%

```{r}
t.test(fruit.win, choc.win)
```

Two functions to use `sort()` and `order()`

```{r}
play <- c(2,1,5,3)

order(play)

```

```{r}
play <- c(2,1,5,3)

sort(play)

```


```{r}
l <- c("c", "a", "b")

sort(l)
```

```{r}
l <- c("c", "a", "b")

order(l)
```


Q13. What are the five least liked candy types in this set?
#Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, Jawbusters

```{r}
inds <- order(candy$winpercent)
head(candy[inds,])
```


Q14. What are the top 5 all time favorite candy types out of this set?
#Reese's Peanut Butter cup, Reese's Miniatures, Twix, Kit Kat, Snickers


```{r}
inds <- order(desc(candy$winpercent))
head(candy[inds,])
```


Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
ggplot(candy) +
  aes(x=winpercent, y= rownames(candy)) +
  geom_col()
```

Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) +
  aes(x=winpercent, y= reorder (rownames(candy), winpercent)) +
  geom_col()
```

I want to define my own colors
I am start with a place-holder vector of colors all "black"


```{r}
mycols <- rep("black", nrow(candy))
mycols[as.logical(candy$chocolate)] <- "chocolate"
mycols[as.logical(candy$bar)] <- "brown"
mycols[as.logical(candy$fruity)] <- "pink"
mycols[rownames(candy) == "Milky Way"] <- "yellow"
```

```{r}
ggplot(candy) +
  aes(x=winpercent, y= reorder (rownames(candy), winpercent)) +
  geom_col(fill = mycols)
```




```{r}
ggplot(candy) +
  aes(winpercent,pricepercent, label = rownames(candy)) +
  geom_point(col= mycols) +
  geom_text(col= mycols)
```



```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=mycols)

```


```{r}
library(corrplot)
```


```{r}
cij <- cor(candy)
corrplot(cij)
```




#Principal Component Analysis 

```{r}
pca <- prcomp(candy, scale= TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2], col=mycols, pch= 16)
```


```{r}
pca$rotation[,1]
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

