---
title: "Class 11"
author: "Tim"
format: pdf
---

Today before delving into structure prediction with AlphaFold we will finish off our previous lab10v"comparative structure analysis" section 

```{r}
library(bio3d)

id <- "1ake_A"

aa <- get.seq(id)
aa
```


```{r}
b <- blast.pdb(aa)

```


```{r}
attributes(b)
```

```{r}
head(b$hit.tbl)
```

```{r}
hits <- plot(b)
```


```{r}
attributes(hits)
```

Top hits that we 

```{r}
hits$pdb.id
```


```{r}
#download releated PDB files 
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip = TRUE)
```


I have now found and downloaded all the ADK structures in the PDB database but viewing them is difficult as they need to be aligned and superpossed.

I am going to install BiocManager package from CRAN. Then I can use `BiocManager::install()`vto install any bioconductor package. 

```{r}
BiocManager::install()

```

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```


```{r}
pc <- pca(pdbs)
plot(pc)
```

```{r}
plot(pc, pc.axes = c(1:2))
```

to examine in more details what PC1 (or any Pc) is capturing here we can plot the loading or make a wee movie (trajectory) of moving along PC1


```{r}
mktrj(pc, pc=1, file = "pc1.pdb")
```


